---
title: "Correlations + Mixed models (Igf & animal welfare)"
author: "Anja Eggert & Anne-Marie Galow"
date: "`r Sys.Date()`" 
editor: visual
code-fold: false
toc: true
format: html
self-contained: true
execute:
  error: true   #  ignore error of code chunk and continue compiling html
#bibliography: references.bib
---

# R Libraries

```{r, libraries, warning=FALSE, message=FALSE, echo=TRUE}
library(tidyverse)     # tidy universe
library(kableExtra)    # html-table
library(patchwork)     # combine plots
library(lmerTest)      # mixed model
library(car)           # ANOVA
library(emmeans)       # post-hoc
library(performance)   # model performance
library(Hmisc)         # Computes correlation and p-value matrix
library(viridis)       # colour scale
```

```{r, my_theme, echo = FALSE}
my_theme = theme_classic() +
  theme(text = element_text(family = "sans"),
        plot.title   = element_text(face = "plain", size = 16),
        axis.title   = element_text(face = "bold",  size = 18),
        legend.title = element_text(face = "plain", size = 16),  
        legend.text  = element_text(face = "plain", size = 16),
        axis.text    = element_text(face = "bold",  size = 16, 
                                    angle = 0, vjust = 0.5),
        panel.grid.major.x      = element_blank(),
        panel.grid.major.y      = element_line(color = "gray", 
                                               linewidth = 0.5, 
                                               linetype = "dotted"),
        panel.grid.minor        = element_blank(),
        axis.ticks.length       = unit(5, "pt"),
        axis.minor.ticks.length = rel(0.5))

my_theme_cor = theme_minimal() +
  theme(text = element_text(family = "sans"),
        plot.title   = element_text(face = "plain", size = 16),
        axis.title   = element_text(face = "bold",  size = 16),
        legend.title = element_text(face = "plain", size = 16),  
        legend.text  = element_text(face = "plain", size = 16),
        axis.text    = element_text(face = "bold",  size = 14, 
                                    angle = 0, vjust = 0.5),
        panel.grid.major        = element_blank(),
        panel.grid.minor        = element_blank(),
        axis.ticks.length       = unit(5, "pt"),
        axis.minor.ticks.length = rel(0.5))
```

```{r, seed}
set.seed(1989)
```

# Data

## Read data

Data processing done in file "igf-biomarker-summary-stats.qmd".

```{r, read-data}
load("./data/data-processed.RData")
```

# Correlations of parameters

## All pairwise correlations

```{r}
dat.cor.all <- dat.w |> 
  dplyr::select(where(is.numeric))
```

Calculate correlations and p values:

```{r}
cor.all <- rcorr(as.matrix(dat.cor.all))

cor.all_matrix  <- round(cor.all$r, 2)
pval.all_matrix <- round(cor.all$P, 3)
```

```{r}
cor.all_matrix |> 
  kable(caption = "Person correlation coefficients") |>
  kable_styling(bootstrap_options = c("striped", "hover"), 
                font_size = 12) |> 
  scroll_box()

pval.all_matrix |> 
  kable(caption = "p values of Person correlation coefficients") |>
  kable_styling(bootstrap_options = c("striped", "hover"), 
                font_size = 12) |> 
  scroll_box()
```

## Selected correlations

Select columns and merge data frames:

```{r}
dat.cor.sel <- dat.w |> 
  dplyr::select(mean.bodyweight, prop.lbw, igf2.ser.105dpc, igf2.ser.8dpp) |> 
  mutate(prop.lbw = prop.lbw/100) |> 
  rename(`Piglet birth weight` = mean.bodyweight,
         `Proportion LBW`      = prop.lbw,
         `IGF2 105dpc`         = igf2.ser.105dpc,
         `IGF2 8dpp`           = igf2.ser.8dpp)
```

Calculate correlations and p values:

```{r}
cor.sel <- rcorr(as.matrix(dat.cor.sel))

cor.sel_matrix  <- as_tibble(cor.sel$r, rownames = "Var1")
pval.sel_matrix <- as_tibble(cor.sel$P, rownames = "Var1")
```

```{r}
cor.sel_matrix |> 
  kable(caption = "Person correlation coefficients") |>
  kable_styling(bootstrap_options = c("striped", "hover"), 
                font_size = 12) |> 
  scroll_box()

pval.sel_matrix |> 
  kable(caption = "p values of Person correlation coefficients") |>
  kable_styling(bootstrap_options = c("striped", "hover"), 
                font_size = 12) |> 
  scroll_box()
```

Heatmap:

```{r}
# Convert matrices into long format for ggplot
cor_long <- pivot_longer(cor.sel_matrix,
                         cols = !Var1,
                         names_to = "Var2",
                         values_to = "cor.val")

p_long   <- pivot_longer(pval.sel_matrix,
                         cols = !Var1,
                         names_to = "Var2",
                         values_to = "p.val")

# Merge correlation values and p-values
heatmap_data <- cor_long |> 
  left_join(p_long, by = c("Var1", "Var2")) |> 
  # this makes it to characters
  mutate(cor.val = sprintf("%.2f", cor.val),
         p.val = sprintf("%.3f", p.val)) |> 
  mutate(cor.val = case_when(cor.val == "1.00" ~ NA_character_, 
                             TRUE ~ cor.val)) |> 
  mutate(p.val = case_when(p.val == "0.000" ~ "<0.001",
                           is.na(p.val) ~ NA_character_, 
                           TRUE ~ paste0("=",p.val))) |>
  mutate(label = case_when(is.na(cor.val) ~ NA_character_, 
                           TRUE ~ paste0(cor.val, "\n(p", p.val, ")")))
```

```{r}
# Create the heatmap
plot.corr <- heatmap_data |> 
  ggplot(aes(Var1, Var2, fill = as.numeric(cor.val))) +
  geom_tile(color = "white") +
  scale_fill_viridis(limits = c(-1, 1), guide="none") +
  # Overlay correlation & p-values
  geom_text(aes(label = label), color = "white", size = 4) + 
  my_theme_cor +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "",
       x = "",
       y = "")
```

```{r, fig.height=8, fig.width=8, warning=FALSE}
plot.corr
```

# Statistical model: general design

To correctly account for the nested data structure in the analysis using a linear mixed model with the `lmerTest` package in R, we need to account for the hierarchical structure in the data:

-   Husbandry Level (conventional/ecological): Each sow is housed in either of the two facilities.
-   Sow Level: Each sow gave birth 1, 2 or 3 times, so these events are nested within sows.
-   Time Point Level: Measurements were taken before and after each birth, introducing repeated measures within each pregnancy.
-   control for age of the sows

This type of nested structure is best represented by a model with random effects at the different levels to account for the dependencies within each level. The random effects allow for variation within subjects and

```{r}
contr = lmerControl(optimizer   = "bobyqa",
                    optCtrl     = list(maxfun = 10000000),
                    calc.derivs = FALSE)
```

# Cortisol serum

## Model

```{r}
hist(dat.l |> 
       dplyr::filter(parameter == "cort.ser") |> 
       droplevels() |>
       drop_na(value) |> 
       pull("value"), 
     breaks = 30)

hist(log(dat.l |> 
         dplyr::filter(parameter == "cort.ser") |> 
         droplevels() |>
         drop_na(value) |> 
         pull("value")),
     breaks = 30)
```

Singularity issues if nested structure defined like `(1 | sow/litter.no)`.

```{r}
mod.cort1 <- lmerTest::lmer(log(value) ~ 
                              husbandry +
                              time +
                              # interaction term
                              husbandry : time +
                              # random intercept for sows and for each litter within sow
                              (1 | sow/litter.no),
                            data    = dat.l |> 
                              dplyr::filter(parameter == "cort.ser") |> 
                              droplevels() |>
                              drop_na(value),
                            REML    = TRUE,
                            control = contr)
```

```{r}
summary(mod.cort1)
```

```{r}
round(drop1(mod.cort1, test = 'Chisq'), 3)
```

Without significant interactions, choose `type = 2`. With significant interactions, choose `type = 3`.

```{r}
car::Anova(mod.cort1,
           test.statistic = "Chisq",
           type = 2)
```

```{r}
car::Anova(mod.cort1,
           test.statistic = "F",
           type = 2)
```

## Model diagnostics

```{r, fig.height=30, fig.width=20, warning=FALSE}
performance::check_model(mod.cort1)
```

## Emmeans & Effect sizes

Most popular effect-size measure is probably Cohen’s d, which is defined as the observed difference, divided by the population SD; and obviously Cohen effect sizes are close cousins of pairwise differences. They are available via the `eff_size()` function of the `emmeans` package.

### Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.cort1,
        pairwise ~ husbandry, 
        data    = dat.l |> 
          dplyr::filter(parameter == "cort.ser") |> 
          droplevels() |>
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.cort1),      # Residual standard deviation from model
         edf = df.residual(mod.cort1))  # Residual degrees of freedom
```

### Time point

Emmeans:

```{r}
emm <- emmeans(mod.cort1,
        pairwise ~ time, 
        data    = dat.l |> 
          dplyr::filter(parameter == "cort.ser") |> 
          droplevels() |>
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.cort1),
         edf = df.residual(mod.cort1))
```

### Husbandry:Time point

Emmeans:

```{r}
emm <- emmeans(mod.cort1,
        pairwise ~ husbandry|time, 
        data    = dat.l |> 
          dplyr::filter(parameter == "cort.ser") |> 
          droplevels() |>
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.cort1),
         edf = df.residual(mod.cort1))
```

### Time point:Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.cort1,
        pairwise ~ time|husbandry, 
        data    = dat.l |> 
          dplyr::filter(parameter == "cort.ser") |> 
          droplevels() |>
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.cort1),
         edf = df.residual(mod.cort1))
```

## Plot

```{r}
plot.cort1 <- dat.l  |> 
  dplyr::filter(parameter == "cort.ser") |> 
  droplevels() |>
  drop_na(value) |> 
  # make plot
  mutate(jit = jitter(as.numeric(time), 0.3)) |>  
  ggplot(aes(y   = value)) +
  geom_boxplot(aes(x   = time, 
                   fill = husbandry),
               col = "black",
               outlier.shape = NA, width = 0.5) +
  geom_jitter(aes(x   = time,
                  fill = husbandry),
              col = "black",
              position = position_jitterdodge(jitter.width  = 0.15,
                                              dodge.width   = 0.5),
              size = 2, alpha = 0.7,
              show.legend = FALSE) +
  scale_fill_manual(labels = c("Conventional",
                               "Ecological"),
                    values = c("#FFA040", "#008000")) +
  scale_y_continuous(lim = c(0, 60), 
                     breaks = seq(0, 60, 20),
                     minor_breaks = seq(0, 60, by = 2) ) +
  scale_x_discrete(labels= c("105dpc", "8dpp")) +
  labs(x = "",
       y = "Cortisol (serum) [ng/ml]",
       fill = "Husbandry") +
  my_theme +
  theme(legend.position = "top") +
  guides(y = guide_axis(minor.ticks = TRUE),
         fill = guide_legend(override.aes = list(linetype = 0, size  = 5)))
```

```{r, fig.height=8, fig.width=8, warning=FALSE}
plot.cort1
```

# Cortisol saliva

## Model

```{r}
hist(dat.l |> 
       dplyr::filter(parameter == "cort.sal") |> 
       droplevels() |>
       drop_na(value) |> 
       pull("value"), 
     breaks = 30)

hist(log(dat.l |> 
         dplyr::filter(parameter == "cort.sal") |> 
         droplevels() |>
         drop_na(value) |> 
         pull("value")),
     breaks = 30)
```

Singularity issues if nested structure defined like `(1 | sow/litter.no)`.

```{r}
mod.cort2 <- lmerTest::lmer(log(value) ~ 
                              husbandry +
                              time +
                              # interaction term
                              husbandry : time +
                              # random intercept for sows and for each litter within sow
                              (1 | sow/litter.no),
                            data    = dat.l |> 
                              dplyr::filter(parameter == "cort.sal") |> 
                              droplevels() |>
                              drop_na(value),
                            REML    = TRUE,
                            control = contr)
```

```{r}
summary(mod.cort2)
```

```{r}
round(drop1(mod.cort2, test = 'Chisq'), 3)
```

Without significant interactions, choose `type = 2`. With significant interactions, choose `type = 3`.

```{r}
car::Anova(mod.cort2,
           test.statistic = "Chisq",
           type = 2)
```

```{r}
car::Anova(mod.cort2,
           test.statistic = "F",
           type = 2)
```

## Model diagnostics

```{r, fig.height=30, fig.width=20, warning=FALSE}
performance::check_model(mod.cort2)
```

## Emmeans & Effect sizes

### Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.cort2,
        pairwise ~ husbandry, 
        data    = dat.l |> 
          dplyr::filter(parameter == "cort.sal") |> 
          droplevels() |>
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.cort2),
         edf = df.residual(mod.cort2))
```

### Time point

Emmeans:

```{r}
emm <- emmeans(mod.cort2,
        pairwise ~ time, 
        data    = dat.l |> 
          dplyr::filter(parameter == "cort.sal") |> 
          droplevels() |>
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.cort2),
         edf = df.residual(mod.cort2))
```

### Husbandry:Time point

Emmeans:

```{r}
emm <- emmeans(mod.cort2,
        pairwise ~ husbandry|time, 
        data    = dat.l |> 
          dplyr::filter(parameter == "cort.sal") |> 
          droplevels() |>
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.cort2),
         edf = df.residual(mod.cort2))
```

### Time point:Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.cort2,
        pairwise ~ time|husbandry, 
        data    = dat.l |> 
          dplyr::filter(parameter == "cort.sal") |> 
          droplevels() |>
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.cort2),
         edf = df.residual(mod.cort2))
```

## Plot

```{r}
plot.cort2 <- dat.l  |> 
  dplyr::filter(parameter == "cort.sal") |> 
  droplevels() |>
  drop_na(value) |> 
  # make plot
  mutate(jit = jitter(as.numeric(time), 0.3)) |>  
  ggplot(aes(y   = value)) +
  geom_boxplot(aes(x   = time, 
                   fill = husbandry),
               col = "black",
               outlier.shape = NA, width = 0.5) +
  geom_jitter(aes(x   = time,
                  fill = husbandry),
              col = "black",
              position = position_jitterdodge(jitter.width  = 0.15,
                                              dodge.width   = 0.5),
              size = 2, alpha = 0.7,
              show.legend = FALSE) +
  scale_fill_manual(labels = c("Conventional",
                               "Ecological"),
                    values = c("#FFA040", "#008000")) +
  scale_y_continuous(lim = c(0, 30), 
                     breaks = seq(0, 30, 10),
                     minor_breaks = seq(0, 30, by = 2) ) +
  scale_x_discrete(labels= c("105dpc", "8dpp")) +
  labs(x = "",
       y = "Cortisol (saliva) [ng/mg protein]",
       fill = "Husbandry") +
  my_theme +
  theme(legend.position = "top") +
  guides(y = guide_axis(minor.ticks = TRUE),
         fill = guide_legend(override.aes = list(linetype = 0, size  = 5)))
```

```{r, fig.height=8, fig.width=8, warning=FALSE}
plot.cort2
```

# IGF bioactivity

## Model

```{r}
hist(dat.l |> 
       dplyr::filter(parameter == "bioact.ser") |> 
       droplevels() |>
       drop_na(value) |> 
       pull("value"), 
     breaks = 30)

hist(log(dat.l |> 
         dplyr::filter(parameter == "bioact.ser") |> 
         droplevels() |>
         drop_na(value) |> 
         pull("value")),
     breaks = 30)
```

Singularity issues if nested structure defined like `(1 | sow/litter.no)`.

```{r}
mod.bioact <- lmerTest::lmer(log(value) ~ 
                               husbandry +
                               time +
                               # interaction term
                               husbandry : time +
                               # random intercept for sows and for each litter within sow
                               (1 | sow/litter.no),
                             data    = dat.l |> 
                               dplyr::filter(parameter == "bioact.ser") |> 
                               drop_na(value) |> 
                               dplyr::filter(time != "30dpc") |> 
                               droplevels(),
                            REML    = TRUE,
                            control = contr)
```

```{r}
summary(mod.bioact)
```

```{r}
round(drop1(mod.bioact, test = 'Chisq'), 3)
```

Without significant interactions, choose `type = 2`. With significant interactions, choose `type = 3`.

```{r}
car::Anova(mod.bioact,
           test.statistic = "Chisq",
           type = 2)
```

```{r}
car::Anova(mod.bioact,
           test.statistic = "F",
           type = 2)
```

## Model diagnostics

```{r, fig.height=30, fig.width=20, warning=FALSE}
performance::check_model(mod.bioact)
```

## Emmeans & Effect sizes

### Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.bioact,
        pairwise ~ husbandry, 
        data    = dat.l |> 
          dplyr::filter(parameter == "bioact.ser") |> 
          drop_na(value) |> 
          dplyr::filter(time != "30dpc") |> 
          droplevels(), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.bioact),
         edf = df.residual(mod.bioact))
```

### Time point

Emmeans:

```{r}
emm <- emmeans(mod.cort2,
        pairwise ~ time, 
        data    = dat.l |> 
          dplyr::filter(parameter == "bioact.ser") |> 
          drop_na(value) |> 
          dplyr::filter(time != "30dpc") |> 
          droplevels(), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.bioact),
         edf = df.residual(mod.bioact))
```

### Husbandry:Time point

Emmeans:

```{r}
emm <- emmeans(mod.bioact,
        pairwise ~ husbandry|time, 
        data    = dat.l |> 
          dplyr::filter(parameter == "bioact.ser") |> 
          drop_na(value) |> 
          dplyr::filter(time != "30dpc") |> 
          droplevels(), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.bioact),
         edf = df.residual(mod.bioact))
```

### Time point:Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.bioact,
        pairwise ~ time|husbandry, 
        data    = dat.l |> 
          dplyr::filter(parameter == "bioact.ser") |> 
          drop_na(value) |> 
          dplyr::filter(time != "30dpc") |> 
          droplevels(), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.bioact),
         edf = df.residual(mod.bioact))
```

## Plot

```{r}
plot.bioact <- dat.l  |> 
  dplyr::filter(parameter == "bioact.ser") |> 
  drop_na(value) |> 
  dplyr::filter(time != "30dpc") |> 
  droplevels() |>
  # plot
  mutate(jit = jitter(as.numeric(time), 0.3)) |>  
  ggplot(aes(y   = value)) +
  geom_boxplot(aes(x   = time, 
                   fill = husbandry),
               col = "black",
               outlier.shape = NA, width = 0.5) +
  geom_jitter(aes(x   = time,
                  fill = husbandry),
              col = "black",
              position = position_jitterdodge(jitter.width  = 0.15,
                                              dodge.width   = 0.5),
              size = 2, alpha = 0.7,
              show.legend = FALSE) +
  scale_fill_manual(labels = c("Conventional",
                               "Ecological"),
                    values = c("#FFA040", "#008000")) +
  scale_y_continuous(lim = c(0, 600), 
                     breaks = seq(0, 600, 200),
                     minor_breaks = seq(0, 600, by = 50) ) +
  scale_x_discrete(labels= c("105dpc", "8dpp")) +
  labs(x = "",
       y = "IGF bioactivity [ng/ml]",
       fill = "Husbandry") +
  my_theme +
  theme(legend.position = "top") +
  guides(y = guide_axis(minor.ticks = TRUE),
         fill = guide_legend(override.aes = list(linetype = 0, size  = 5)))
```

```{r, fig.height=8, fig.width=8, warning=FALSE}
plot.bioact
```

# IGF1 (serum)

## Model

```{r}
hist(dat.l |> 
       dplyr::filter(parameter == "igf1.ser") |> 
       droplevels() |> 
       drop_na(value) |> 
       pull("value"), 
     breaks = 30)

hist(log(dat.l |> 
       dplyr::filter(parameter == "igf1.ser") |> 
       droplevels() |> 
       drop_na(value) |> 
       pull("value")),
     breaks = 30)
```

Singularity issues if nested structure defined like `(1 | sow/litter.no)`.

```{r}
mod.igf1 <- lmerTest::lmer(log(value) ~ 
                              husbandry +
                              time +
                              # interaction term
                              husbandry : time +
                              # random intercept for sows and for each litter within sow
                              (1 | sow/litter.no),
                            data    = dat.l |> 
                              dplyr::filter(parameter == "igf1.ser") |> 
                              droplevels() |>
                              drop_na(value),
                            REML    = TRUE,
                            control = contr)
```

```{r}
summary(mod.igf1)
```

```{r}
round(drop1(mod.igf1, test = 'Chisq'), 3)
```

Without significant interactions, choose `type = 2`. With significant interactions, choose `type = 3`.

```{r}
car::Anova(mod.igf1,
           test.statistic = "Chisq",
           type = 2)
```

```{r}
car::Anova(mod.igf1,
           test.statistic = "F",
           type = 2)
```

## Model diagnostics

```{r, fig.height=30, fig.width=20, warning=FALSE}
performance::check_model(mod.igf1)
```

## Emmeans & Effect sizes

### Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.igf1,
        pairwise ~ husbandry, 
        data    = dat.l |> 
          dplyr::filter(parameter == "igf1.ser") |> 
          droplevels() |>
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.igf1),
         edf = df.residual(mod.igf1))
```

### Time point

Emmeans:

```{r}
emm <- emmeans(mod.igf1,
        pairwise ~ time, 
        data    = dat.l |> 
          dplyr::filter(parameter == "igf1.ser") |> 
          droplevels() |>
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.igf1),
         edf = df.residual(mod.igf1))
```

### Husbandry:Time point

Emmeans:

```{r}
emm <- emmeans(mod.igf1,
        pairwise ~ husbandry|time, 
        data    = dat.l |> 
          dplyr::filter(parameter == "igf1.ser") |> 
          droplevels() |>
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.igf1),
         edf = df.residual(mod.igf1))
```

### Time point:Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.igf1,
        pairwise ~ time|husbandry, 
        data    = dat.l |> 
          dplyr::filter(parameter == "igf1.ser") |> 
          droplevels() |>
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.igf1),
         edf = df.residual(mod.igf1))
```

## Plot

```{r}
plot.igf1 <- dat.l  |> 
  dplyr::filter(parameter == "igf1.ser") |> 
  droplevels() |>
  drop_na(value) |> 
  # plot
  mutate(jit = jitter(as.numeric(time), 0.3)) |>  
  ggplot(aes(y   = value)) +
  geom_boxplot(aes(x   = time, 
                   fill = husbandry),
               col = "black",
               outlier.shape = NA, width = 0.5) +
  geom_jitter(aes(x   = time,
                  fill = husbandry),
              col = "black",
              position = position_jitterdodge(jitter.width  = 0.15,
                                              dodge.width   = 0.5),
              size = 2, alpha = 0.7,
              show.legend = FALSE) +
  scale_fill_manual(labels = c("Conventional",
                               "Ecological"),
                    values = c("#FFA040", "#008000")) +
  scale_y_continuous(lim = c(0, 500), 
                     breaks = seq(0, 500, 100),
                     minor_breaks = seq(0, 500, by = 20) ) +
  scale_x_discrete(labels= c("105dpc", "8dpp")) +
  labs(x = "",
       y = "IGF1 (serum) [ng/ml]",
       fill = "Husbandry") +
  my_theme +
  theme(legend.position = "top") +
  guides(y = guide_axis(minor.ticks = TRUE),
         fill = guide_legend(override.aes = list(linetype = 0, size  = 5)))
```

```{r, fig.height=8, fig.width=8, warning=FALSE}
plot.igf1
```

# IGF2 (serum)

## Model

```{r}
hist(dat.l |> 
       dplyr::filter(parameter == "igf2.ser") |> 
       droplevels() |> 
       drop_na(value) |> 
       pull("value"), 
     breaks = 30)

hist(log(dat.l |> 
       dplyr::filter(parameter == "igf2.ser") |> 
       droplevels() |>
       drop_na(value) |> 
       pull("value")),
     breaks = 30)
```

Singularity issues if nested structure defined like `(1 | sow/litter.no)`.

```{r}
mod.igf2 <- lmerTest::lmer(log(value) ~ 
                              husbandry +
                              time +
                              # interaction term
                              husbandry : time +
                              # random intercept for sows and for each litter within sow
                              (1 | sow/litter.no),
                            data    = dat.l |> 
                              dplyr::filter(parameter == "igf2.ser") |> 
                              droplevels() |> 
                              drop_na(value),
                            REML    = TRUE,
                            control = contr)
```

```{r}
summary(mod.igf2)
```

```{r}
round(drop1(mod.igf2, test = 'Chisq'), 3)
```

Without significant interactions, choose `type = 2`. With significant interactions, choose `type = 3`.

```{r}
car::Anova(mod.igf2,
           test.statistic = "Chisq",
           type = 2)
```

```{r}
car::Anova(mod.igf2,
           test.statistic = "F",
           type = 2)
```

## Model diagnostics

```{r, fig.height=30, fig.width=20, warning=FALSE}
performance::check_model(mod.igf2)
```

## Emmeans & Effect sizes

### Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.igf2,
        pairwise ~ husbandry, 
        data    = dat.l |> 
          dplyr::filter(parameter == "igf2.ser") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.igf2),
         edf = df.residual(mod.igf2))
```

### Time point

Emmeans:

```{r}
emm <- emmeans(mod.igf2,
        pairwise ~ time, 
        data    = dat.l |> 
          dplyr::filter(parameter == "igf2.ser") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.igf2),
         edf = df.residual(mod.igf2))
```

### Husbandry:Time point

Emmeans:

```{r}
emm <- emmeans(mod.igf2,
        pairwise ~ husbandry|time, 
        data    = dat.l |> 
          dplyr::filter(parameter == "igf2.ser") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.igf2),
         edf = df.residual(mod.igf2))
```

### Time point:Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.igf2,
        pairwise ~ time|husbandry, 
        data    = dat.l |> 
          dplyr::filter(parameter == "igf2.ser") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.igf2),
         edf = df.residual(mod.igf2))
```

## Plot

```{r}
plot.igf2 <- dat.l  |> 
  dplyr::filter(parameter == "igf2.ser") |> 
  droplevels() |> 
  drop_na(value) |> 
  # plot
  mutate(jit = jitter(as.numeric(time), 0.3)) |>  
  ggplot(aes(y   = value)) +
  geom_boxplot(aes(x   = time, 
                   fill = husbandry),
               col = "black",
               outlier.shape = NA, width = 0.5) +
  geom_jitter(aes(x   = time,
                  fill = husbandry),
              col = "black",
              position = position_jitterdodge(jitter.width  = 0.15,
                                              dodge.width   = 0.5),
              size = 2, alpha = 0.7,
              show.legend = FALSE) +
  scale_fill_manual(labels = c("Conventional",
                               "Ecological"),
                    values = c("#FFA040", "#008000")) +
  scale_y_continuous(lim = c(0, 200), 
                     breaks = seq(0, 200, 50),
                     minor_breaks = seq(0, 200, by = 10) ) +
  scale_x_discrete(labels= c("105dpc", "8dpp")) +
  labs(x = "",
       y = "IGF2 (serum) [ng/ml]",
       fill = "Husbandry") +
  guides(y = guide_axis(minor.ticks = TRUE),
         fill = guide_legend(override.aes = list(linetype = 0, size  = 5))) +
  my_theme +
  theme(legend.position = c(0.25, 0.85),
        legend.box = "horizontal")
```

```{r, fig.height=8, fig.width=8, warning=FALSE}
plot.igf2
```

# Combined plots: Figure 3

```{r}
# Combine plots with a designated area for the legend
combined <- (plot.cort1 + 
             plot.cort2 + 
             plot.bioact + 
             plot.igf1) + 
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(face = "bold", size = 20),
        legend.position = "bottom",
        legend.direction = "horizontal")
```

```{r, fig.height=20, fig.width=20, warning=FALSE}
combined
```

```{r, eval = TRUE}
png("./plots/figure3.png",
     width = 300, height = 300, units = "mm",
     pointsize = 10, res = 600)

combined

dev.off()
```

# Combined plots: Figure 4

```{r}
# Combine plots with a designated area for the legend
combined <- (plot.igf2 +
             plot.corr) + 
  #plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(face = "bold", size = 20))
```

```{r, fig.height=20, fig.width=20, warning=FALSE}
combined
```

```{r, eval = TRUE}
png("./plots/figure4.png",
     width = 300, height = 200, units = "mm",
     pointsize = 10, res = 600)

combined

dev.off()
```

# Litter numbers

```{r}
plot.bio.li <- dat.w  |> 
  # plot
  mutate(jit = jitter(as.numeric(litter.no), 0.3)) |>  
  ggplot(aes(y   = bioact.ser.105dpc)) +
  geom_boxplot(aes(x   = litter.no, 
                   fill = husbandry),
               col = "black",
               outlier.shape = NA, width = 0.5) +
  geom_jitter(aes(x   = litter.no,
                  fill = husbandry),
              col = "black",
              position = position_jitterdodge(jitter.width  = 0.15,
                                              dodge.width   = 0.5),
              size = 2, alpha = 0.7,
              show.legend = FALSE) +
  scale_fill_manual(labels = c("Conventional",
                               "Ecological"),
                    values = c("#FFA040", "#008000")) +
  scale_y_continuous(lim = c(0, 600), 
                     breaks = seq(0, 600, 200),
                     minor_breaks = seq(0, 600, by = 50) ) +
  scale_x_discrete(labels= c("1st", "2nd", "3rd")) +
  labs(x = "Litter number",
       y = "IGF bioactivity 105dpc [ng/ml]",
       fill = "Husbandry") +
  my_theme +
  theme(legend.position = "bottom") +
  guides(y = guide_axis(minor.ticks = TRUE),
         fill = guide_legend(override.aes = list(linetype = 0, size  = 5)))
```

```{r}
plot.lbw.li <- dat.w  |> 
  # plot
  mutate(jit = jitter(as.numeric(litter.no), 0.3)) |>  
  ggplot(aes(y   = prop.lbw)) +
  geom_boxplot(aes(x   = litter.no, 
                   fill = husbandry),
               col = "black",
               outlier.shape = NA, width = 0.5) +
  geom_jitter(aes(x   = litter.no,
                  fill = husbandry),
              col = "black",
              position = position_jitterdodge(jitter.width  = 0.15,
                                              dodge.width   = 0.5),
              size = 2, alpha = 0.7,
              show.legend = FALSE) +
  scale_fill_manual(labels = c("Conventional",
                               "Ecological"),
                    values = c("#FFA040", "#008000")) +
  scale_y_continuous(lim = c(0, 100), 
                     breaks = seq(0, 100, 20),
                     minor_breaks = seq(0, 100, by = 5) ) +
  scale_x_discrete(labels= c("1st", "2nd", "3rd")) +
  labs(x = "Litter number",
       y = "Proportion LBW [%]",
       fill = "Husbandry") +
  my_theme +
  theme(legend.position = "bottom") +
  guides(y = guide_axis(minor.ticks = TRUE),
         fill = guide_legend(override.aes = list(linetype = 0, size  = 5)))
```

# Combined plots: Figure 5

```{r}
# Combine plots with a designated area for the legend
combined <- (plot.bio.li + 
             plot.lbw.li) + 
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(face = "bold", size = 20),
        legend.position = "bottom")
```

```{r, fig.height=20, fig.width=20, warning=FALSE}
combined
```

```{r, eval = TRUE}
png("./plots/figure5.png",
     width = 300, height = 200, units = "mm",
     pointsize = 10, res = 600)

combined

dev.off()
```

# IGFBP2 (serum)

## Model

```{r}
hist(dat.l |> 
       dplyr::filter(parameter == "igfbp2.ser") |> 
       droplevels() |> 
       drop_na(value) |> 
       pull("value"), 
     breaks = 30)

hist(log(dat.l |> 
         dplyr::filter(parameter == "igfbp2.ser") |> 
         droplevels() |>
         drop_na(value) |> 
         pull("value")),
     breaks = 30)
```

Singularity issues if nested structure defined like `(1 | sow/litter.no)`.

```{r}
mod.igfbp2 <- lmerTest::lmer(log(value) ~ 
                               husbandry +
                               time +
                               # interaction term
                               husbandry : time +
                               # random intercept for sows and for each litter within sow
                               (1 | sow/litter.no),
                             data    = dat.l |> 
                               dplyr::filter(parameter == "igfbp2.ser") |> 
                               droplevels() |> 
                               drop_na(value),
                             REML    = TRUE,
                             control = contr)
```

```{r}
summary(mod.igfbp2)
```

```{r}
round(drop1(mod.igfbp2, test = 'Chisq'), 3)
```

Without significant interactions, choose `type = 2`. With significant interactions, choose `type = 3`.

```{r}
car::Anova(mod.igfbp2,
           test.statistic = "Chisq",
           type = 2)
```

```{r}
car::Anova(mod.igfbp2,
           test.statistic = "F",
           type = 2)
```

## Model diagnostics

```{r, fig.height=30, fig.width=20, warning=FALSE}
performance::check_model(mod.igfbp2)
```

## Emmeans & Effect sizes

### Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.igfbp2,
        pairwise ~ husbandry, 
        data    = dat.l |> 
          dplyr::filter(parameter == "igfbp2.ser") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.igfbp2),
         edf = df.residual(mod.igfbp2))
```

### Time point

Emmeans:

```{r}
emm <- emmeans(mod.igfbp2,
        pairwise ~ time, 
        data    = dat.l |> 
          dplyr::filter(parameter == "igfbp2.ser") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.igfbp2),
         edf = df.residual(mod.igfbp2))
```

### Husbandry:Time point

Emmeans:

```{r}
emm <- emmeans(mod.igfbp2,
        pairwise ~ husbandry|time, 
        data    = dat.l |> 
          dplyr::filter(parameter == "igfbp2.ser") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.igfbp2),
         edf = df.residual(mod.igfbp2))
```

### Time point:Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.igfbp2,
        pairwise ~ time|husbandry, 
        data    = dat.l |> 
          dplyr::filter(parameter == "igfbp2.ser") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.igfbp2),
         edf = df.residual(mod.igfbp2))
```

## Plot

```{r}
plot.igfbp2 <- dat.l  |> 
  dplyr::filter(parameter == "igfbp2.ser") |> 
  droplevels() |> 
  drop_na(value) |> 
  # plot
  mutate(jit = jitter(as.numeric(time), 0.3)) |>  
  ggplot(aes(y   = value)) +
  geom_boxplot(aes(x   = time, 
                   fill = husbandry),
               col = "black",
               outlier.shape = NA, width = 0.5) +
  geom_jitter(aes(x   = time,
                  fill = husbandry),
              col = "black",
              position = position_jitterdodge(jitter.width  = 0.15,
                                              dodge.width   = 0.5),
              size = 2, alpha = 0.7,
              show.legend = FALSE) +
  scale_fill_manual(labels = c("Conventional",
                               "Ecological"),
                    values = c("#FFA040", "#008000")) +
  scale_y_continuous(lim = c(0, 1250), 
                     breaks = seq(0, 1250, 250),
                     minor_breaks = seq(0, 1250, by = 100) ) +
  scale_x_discrete(labels= c("105dpc", "8dpp")) +
  labs(x = "",
       y = "IGF BP2 (serum) [ng/ml]",
       fill = "Husbandry") +
  my_theme +
  theme(legend.position = "top") +
  guides(y = guide_axis(minor.ticks = TRUE),
         fill = guide_legend(override.aes = list(linetype = 0, size  = 5)))
```

```{r, fig.height=8, fig.width=8, warning=FALSE}
plot.igfbp2
```

# IGFBP3 (serum)

## Model

```{r}
hist(dat.l |> 
       dplyr::filter(parameter == "igfbp3.ser") |> 
       droplevels() |> 
       drop_na(value) |> 
       pull("value"), 
     breaks = 30)

hist(log(dat.l |> 
         dplyr::filter(parameter == "igfbp3.ser") |> 
         droplevels() |> 
         drop_na(value) |> 
         pull("value")),
     breaks = 30)
```

Singularity issues if nested structure defined like `(1 | sow/litter.no)`.

```{r}
mod.igfbp3 <- lmerTest::lmer(log(value) ~ 
                               husbandry +
                               time +
                               # interaction term
                               husbandry : time +
                               # random intercept for sows and for each litter within sow
                               (1 | sow/litter.no),
                             data    = dat.l |> 
                               dplyr::filter(parameter == "igfbp3.ser") |> 
                               droplevels() |> 
                               drop_na(value),
                             REML    = TRUE,
                             control = contr)
```

```{r}
summary(mod.igfbp3)
```

```{r}
round(drop1(mod.igfbp3, test = 'Chisq'), 3)
```

Without significant interactions, choose `type = 2`. With significant interactions, choose `type = 3`.

```{r}
car::Anova(mod.igfbp3,
           test.statistic = "Chisq",
           type = 2)
```

```{r}
car::Anova(mod.igfbp3,
           test.statistic = "F",
           type = 2)
```

## Model diagnostics

```{r, fig.height=30, fig.width=20, warning=FALSE}
performance::check_model(mod.igfbp3)
```

## Emmeans & Effect sizes

### Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.igfbp3,
        pairwise ~ husbandry, 
        data    = dat.l |> 
          dplyr::filter(parameter == "igfbp3.ser") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.igfbp3),
         edf = df.residual(mod.igfbp3))
```

### Time point

Emmeans:

```{r}
emm <- emmeans(mod.igfbp3,
        pairwise ~ time, 
        data    = dat.l |> 
          dplyr::filter(parameter == "igfbp3.ser") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.igfbp3),
         edf = df.residual(mod.igfbp3))
```

### Husbandry:Time point

Emmeans:

```{r}
emm <- emmeans(mod.igfbp3,
        pairwise ~ husbandry|time, 
        data    = dat.l |> 
          dplyr::filter(parameter == "igfbp3.ser") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.igfbp3),
         edf = df.residual(mod.igfbp3))
```

### Time point:Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.igfbp3,
        pairwise ~ time|husbandry, 
        data    = dat.l |> 
          dplyr::filter(parameter == "igfbp3.ser") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.igfbp3),
         edf = df.residual(mod.igfbp3))
```

## Plot

```{r}
plot.igfbp3 <- dat.l  |> 
  dplyr::filter(parameter == "igfbp3.ser") |> 
  droplevels() |> 
  drop_na(value) |> 
  # plot
  mutate(jit = jitter(as.numeric(time), 0.3)) |>  
  ggplot(aes(y   = value)) +
  geom_boxplot(aes(x   = time, 
                   fill = husbandry),
               col = "black",
               outlier.shape = NA, width = 0.5) +
  geom_jitter(aes(x   = time,
                  fill = husbandry),
              col = "black",
              position = position_jitterdodge(jitter.width  = 0.15,
                                              dodge.width   = 0.5),
              size = 2, alpha = 0.7,
              show.legend = FALSE) +
  scale_fill_manual(labels = c("Conventional",
                               "Ecological"),
                    values = c("#FFA040", "#008000")) +
  scale_y_continuous(lim = c(0, 6500), 
                     breaks = seq(0, 6500, 2000),
                     minor_breaks = seq(0, 6500, by = 200) ) +
  scale_x_discrete(labels= c("105dpc", "8dpp")) +
  labs(x = "",
       y = "IGF BP3 (serum) [ng/ml]",
       fill = "Husbandry") +
  my_theme +
  theme(legend.position = "top") +
  guides(y = guide_axis(minor.ticks = TRUE),
         fill = guide_legend(override.aes = list(linetype = 0, size  = 5)))
```

```{r, fig.height=8, fig.width=8, warning=FALSE}
plot.igfbp3
```

# IGFBP2/IGFBP3 (serum)

## Model

```{r}
hist(dat.l |> 
       dplyr::filter(parameter == "igfbp23.ser") |> 
       droplevels() |> 
       drop_na(value) |> 
       pull("value"), 
     breaks = 30)

hist(log(dat.l |> 
         dplyr::filter(parameter == "igfbp23.ser") |> 
         droplevels() |>
         drop_na(value) |> 
         pull("value")),
     breaks = 30)
```

Singularity issues if nested structure defined like `(1 | sow/litter.no)`.

```{r}
mod.igfbp23 <- lmerTest::lmer(log(value) ~ 
                               husbandry +
                               time +
                               # interaction term
                               husbandry : time +
                               # random intercept for sows and for each litter within sow
                               (1 | sow/litter.no),
                             data    = dat.l |> 
                               dplyr::filter(parameter == "igfbp23.ser") |> 
                               droplevels() |>
                               drop_na(value),
                             REML    = TRUE,
                             control = contr)
```

```{r}
summary(mod.igfbp23)
```

```{r}
round(drop1(mod.igfbp23, test = 'Chisq'), 3)
```

Without significant interactions, choose `type = 2`. With significant interactions, choose `type = 3`.

```{r}
car::Anova(mod.igfbp23,
           test.statistic = "Chisq",
           type = 2)
```

```{r}
car::Anova(mod.igfbp23,
           test.statistic = "F",
           type = 2)
```

## Model diagnostics

```{r, fig.height=30, fig.width=20, warning=FALSE}
performance::check_model(mod.igfbp23)
```

## Emmeans & Effect sizes

### Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.igfbp23,
        pairwise ~ husbandry, 
        data    = dat.l |> 
          dplyr::filter(parameter == "igfbp23.ser") |> 
          droplevels() |>
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.igfbp23),
         edf = df.residual(mod.igfbp23))
```

### Time point

Emmeans:

```{r}
emm <- emmeans(mod.igfbp23,
        pairwise ~ time, 
        data    = dat.l |> 
          dplyr::filter(parameter == "igfbp23.ser") |> 
          droplevels() |>
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.igfbp23),
         edf = df.residual(mod.igfbp23))
```

### Husbandry:Time point

Emmeans:

```{r}
emm <- emmeans(mod.igfbp23,
        pairwise ~ husbandry|time, 
        data    = dat.l |> 
          dplyr::filter(parameter == "igfbp23.ser") |> 
          droplevels() |>
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.igfbp23),
         edf = df.residual(mod.igfbp23))
```

### Time point:Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.igfbp23,
        pairwise ~ time|husbandry, 
        data    = dat.l |> 
          dplyr::filter(parameter == "igfbp23.ser") |> 
          droplevels() |>
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.igfbp23),
         edf = df.residual(mod.igfbp23))
```

## Plot

```{r}
plot.igfbp23 <- dat.l  |> 
  dplyr::filter(parameter == "igfbp23.ser") |> 
  droplevels() |>
  drop_na(value) |> 
  # plot
  mutate(jit = jitter(as.numeric(time), 0.3)) |>  
  ggplot(aes(y   = (value))) +
  geom_boxplot(aes(x   = time, 
                   fill = husbandry),
               col = "black",
               outlier.shape = NA, width = 0.5) +
  geom_jitter(aes(x   = time,
                  fill = husbandry),
              col = "black",
              position = position_jitterdodge(jitter.width  = 0.15,
                                              dodge.width   = 0.5),
              size = 2, alpha = 0.7,
              show.legend = FALSE) +
  scale_fill_manual(labels = c("Conventional",
                               "Ecological"),
                    values = c("#FFA040", "#008000")) +
  scale_y_continuous(lim = c(0, 2.5), 
                     breaks = seq(0, 2.5, 0.5),
                     minor_breaks = seq(0, 2.5, by = 0.1) ) +
  scale_x_discrete(labels= c("105dpc", "8dpp")) +
  labs(x = "",
       y = "Molar ratio of IGFBP2/IGFBP3 (serum)",
       fill = "Husbandry") +
  my_theme +
  theme(legend.position = "top") +
  guides(y = guide_axis(minor.ticks = TRUE),
         fill = guide_legend(override.aes = list(linetype = 0, size  = 5)))
```

```{r, fig.height=8, fig.width=8, warning=FALSE}
plot.igfbp23
```

# Proteolytic activity (serum)

## Model

```{r}
hist(dat.l |> 
       dplyr::filter(parameter == "proteolysis") |> 
       droplevels() |> 
       drop_na(value) |> 
       pull("value"), 
     breaks = 30)

hist(log(dat.l |> 
         dplyr::filter(parameter == "proteolysis") |> 
         droplevels() |> 
         drop_na(value) |> 
         pull("value")),
     breaks = 30)
```

Singularity issues if nested structure defined like `(1 | sow/litter.no)`.

```{r}
mod.prot <- lmerTest::lmer(log(value+1) ~ 
                               husbandry +
                               time +
                               # interaction term
                               husbandry : time +
                               # random intercept for sows and for each litter within sow
                               (1 | sow/litter.no),
                             data    = dat.l |> 
                              dplyr::filter(parameter == "proteolysis") |> 
                              droplevels() |> 
                              drop_na(value),
                             REML    = TRUE,
                             control = contr)
```

```{r}
summary(mod.prot)
```

```{r}
round(drop1(mod.prot, test = 'Chisq'), 3)
```

Without significant interactions, choose `type = 2`. With significant interactions, choose `type = 3`.

```{r}
car::Anova(mod.prot,
           test.statistic = "Chisq",
           type = 2)
```

```{r}
car::Anova(mod.prot,
           test.statistic = "F",
           type = 2)
```

## Model diagnostics

```{r}
performance::check_model(mod.prot)
```

## Emmeans & Effect sizes

### Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.prot,
        pairwise ~ husbandry, 
        data    = dat.l |> 
          dplyr::filter(parameter == "proteolysis") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.prot),
         edf = df.residual(mod.prot))
```

### Time point

Emmeans:

```{r}
emm <- emmeans(mod.prot,
        pairwise ~ time, 
        data    = dat.l |> 
          dplyr::filter(parameter == "proteolysis") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.prot),
         edf = df.residual(mod.prot))
```

### Husbandry:Time point

Emmeans:

```{r}
emm <- emmeans(mod.prot,
        pairwise ~ husbandry|time, 
        data    = dat.l |> 
          dplyr::filter(parameter == "proteolysis") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.prot),
         edf = df.residual(mod.prot))
```

### Time point:Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.prot,
        pairwise ~ time|husbandry, 
        data    = dat.l |> 
          dplyr::filter(parameter == "proteolysis") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.prot),
         edf = df.residual(mod.prot))
```

## Plot

```{r}
plot.prot <- dat.l  |> 
  dplyr::filter(parameter == "proteolysis") |> 
  droplevels() |> 
  drop_na(value) |> 
  # plot
  mutate(jit = jitter(as.numeric(time), 0.3)) |>  
  ggplot(aes(y   = value)) +
  geom_boxplot(aes(x   = time, 
                   fill = husbandry),
               col = "black",
               outlier.shape = NA, width = 0.5) +
  geom_jitter(aes(x   = time,
                  fill = husbandry),
              col = "black",
              position = position_jitterdodge(jitter.width  = 0.15,
                                              dodge.width   = 0.5),
              size = 2, alpha = 0.7,
              show.legend = FALSE) +
  scale_fill_manual(labels = c("Conventional",
                               "Ecological"),
                    values = c("#FFA040", "#008000")) +
  scale_y_continuous(lim = c(0, 21), 
                     breaks = seq(0, 21, 5),
                     minor_breaks = seq(0, 21, by = 1) ) +
  scale_x_discrete(labels= c("105dpc", "8dpp")) +
  labs(x = "",
       y = "Proteolytic activity (serum) [%]",
       fill = "Husbandry") +
  my_theme +
  theme(legend.position = "top") +
  guides(y = guide_axis(minor.ticks = TRUE),
         fill = guide_legend(override.aes = list(linetype = 0, size  = 5)))
```

```{r, fig.height=8, fig.width=8, warning=FALSE}
plot.prot
```

# Stanniocalcin, STC1 (serum)

## Model

```{r}
hist(dat.l |> 
       dplyr::filter(parameter == "stc1") |> 
       droplevels() |> 
       drop_na(value) |> 
       pull("value"), 
     breaks = 30)

hist(log(dat.l |> 
         dplyr::filter(parameter == "stc1") |> 
         droplevels() |> 
         drop_na(value) |> 
         pull("value")),
     breaks = 30)
```

Singularity issues if nested structure defined like `(1 | sow/litter.no)`.

```{r}
mod.stc1 <- lmerTest::lmer(log(value) ~ 
                               husbandry +
                               time +
                               # interaction term
                               husbandry : time +
                               # random intercept for sows and for each litter within sow
                               (1 | sow/litter.no),
                             data    = dat.l |> 
                               dplyr::filter(parameter == "stc1") |> 
                               droplevels() |> 
                               drop_na(value),
                             REML    = TRUE,
                             control = contr)
```

```{r}
summary(mod.stc1)
```

```{r}
round(drop1(mod.stc1, test = 'Chisq'), 3)
```

Without significant interactions, choose `type = 2`. With significant interactions, choose `type = 3`.

```{r}
car::Anova(mod.stc1,
           test.statistic = "Chisq",
           type = 2)
```

```{r}
car::Anova(mod.stc1,
           test.statistic = "F",
           type = 2)
```

## Model diagnostics

```{r, fig.height=30, fig.width=20, warning=FALSE}
performance::check_model(mod.stc1)
```

## Emmeans & Effect sizes

### Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.stc1,
        pairwise ~ husbandry, 
        data    = dat.l |>
          dplyr::filter(parameter == "stc1") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.stc1),
         edf = df.residual(mod.stc1))
```

### Time point

Emmeans:

```{r}
emm <- emmeans(mod.stc1,
        pairwise ~ time, 
        data    = dat.l |>
          dplyr::filter(parameter == "stc1") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.stc1),
         edf = df.residual(mod.stc1))
```

### Husbandry:Time point

Emmeans:

```{r}
emm <- emmeans(mod.stc1,
        pairwise ~ husbandry|time, 
        data    = dat.l |>
          dplyr::filter(parameter == "stc1") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.stc1),
         edf = df.residual(mod.stc1))
```

### Time point:Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.stc1,
        pairwise ~ time|husbandry, 
        data    = dat.l |>
          dplyr::filter(parameter == "stc1") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.stc1),
         edf = df.residual(mod.stc1))
```

## Plot

```{r}
plot.stc1 <- dat.l  |> 
  dplyr::filter(parameter == "stc1") |> 
  droplevels() |> 
  drop_na(value) |> 
  # plot
  mutate(jit = jitter(as.numeric(time), 0.3)) |>  
  ggplot(aes(y   = value)) +
  geom_boxplot(aes(x   = time, 
                   fill = husbandry),
               col = "black",
               outlier.shape = NA, width = 0.5) +
  geom_jitter(aes(x   = time,
                  fill = husbandry),
              col = "black",
              position = position_jitterdodge(jitter.width  = 0.15,
                                              dodge.width   = 0.5),
              size = 2, alpha = 0.7,
              show.legend = FALSE) +
  scale_fill_manual(labels = c("Conventional",
                               "Ecological"),
                    values = c("#FFA040", "#008000")) +
  scale_y_continuous(lim = c(0, 10000), 
                     breaks = seq(0, 10000, 2000),
                     minor_breaks = seq(0, 10000, by = 400) ) +
  scale_x_discrete(labels= c("105dpc", "8dpp")) +
  labs(x = "",
       y = "STC1 (serum) [pg/ml]",
       fill = "Husbandry") +
  my_theme +
  theme(legend.position = "top") +
  guides(y = guide_axis(minor.ticks = TRUE),
         fill = guide_legend(override.aes = list(linetype = 0, size  = 5)))
```

```{r, fig.height=8, fig.width=8, warning=FALSE}
plot.stc1
```

# Combined plots: Figure 6

```{r}
# Combine plots with a designated area for the legend
combined <- (plot.igfbp23 + 
             plot.stc1) + 
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(face = "bold", size = 20),
        legend.position = "bottom")
```

```{r, fig.height=20, fig.width=20, warning=FALSE}
combined
```

```{r, eval = TRUE}
png("./plots/figure6.png",
     width = 300, height = 200, units = "mm",
     pointsize = 10, res = 600)

combined

dev.off()
```

# Calcium (serum)

## Model

```{r}
hist(dat.l |> 
       dplyr::filter(parameter == "calc") |> 
       droplevels() |> 
       drop_na(value) |> 
       pull("value"), 
     breaks = 30)

hist(log(dat.l |> 
         dplyr::filter(parameter == "calc") |> 
         droplevels() |> 
         drop_na(value) |> 
         pull("value")),
     breaks = 30)
```

Singularity issues if nested structure defined like `(1 | sow/litter.no)`.

```{r}
mod.calc <- lmerTest::lmer(log(value) ~ 
                               husbandry +
                               time +
                               # interaction term
                               husbandry : time +
                               # random intercept for sows and for each litter within sow
                               (1 | sow/litter.no),
                             data    = dat.l |> 
                               dplyr::filter(parameter == "calc") |> 
                               dplyr::filter(time != "30dpc") |> 
                               droplevels() |> 
                               drop_na(value),
                             REML    = TRUE,
                             control = contr)
```

```{r}
summary(mod.calc)
```

```{r}
round(drop1(mod.calc, test = 'Chisq'), 3)
```

Without significant interactions, choose `type = 2`. With significant interactions, choose `type = 3`.

```{r}
car::Anova(mod.calc,
           test.statistic = "Chisq",
           type = 2)
```

```{r}
car::Anova(mod.calc,
           test.statistic = "F",
           type = 2)
```

## Model diagnostics

```{r, fig.height=30, fig.width=20, warning=FALSE}
performance::check_model(mod.calc)
```

## Emmeans & Effect sizes

### Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.calc,
        pairwise ~ husbandry, 
        data    = dat.l |> 
          dplyr::filter(parameter == "calc") |> 
          dplyr::filter(time != "30dpc") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.calc),
         edf = df.residual(mod.calc))
```

### Time point

Emmeans:

```{r}
emm <- emmeans(mod.calc,
        pairwise ~ time, 
        data    = dat.l |> 
          dplyr::filter(parameter == "calc") |> 
          dplyr::filter(time != "30dpc") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.calc),
         edf = df.residual(mod.calc))
```

### Husbandry:Time point

Emmeans:

```{r}
emm <- emmeans(mod.calc,
        pairwise ~ husbandry|time, 
        data    = dat.l |> 
          dplyr::filter(parameter == "calc") |> 
          dplyr::filter(time != "30dpc") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.calc),
         edf = df.residual(mod.calc))
```

### Time point:Husbandry

Emmeans:

```{r}
emm <- emmeans(mod.calc,
        pairwise ~ time|husbandry, 
        data    = dat.l |> 
          dplyr::filter(parameter == "calc") |> 
          dplyr::filter(time != "30dpc") |> 
          droplevels() |> 
          drop_na(value), 
        adjust  = "tukey",
        lmer.df = "satterthwaite",
        type    = "response")

emm
```

Effect sizes:

```{r}
eff_size(emm,
         sigma = sigma(mod.calc),
         edf = df.residual(mod.calc))
```

## Plot

```{r}
plot.calc <- dat.l  |> 
  dplyr::filter(parameter == "calc") |> 
  dplyr::filter(time != "30dpc") |> 
  droplevels() |> 
  drop_na(value) |> 
  # plot
  mutate(jit = jitter(as.numeric(time), 0.3)) |>  
  ggplot(aes(y   = value)) +
  geom_boxplot(aes(x   = time, 
                   fill = husbandry),
               col = "black",
               outlier.shape = NA, width = 0.5) +
  geom_jitter(aes(x   = time,
                  fill = husbandry),
              col = "black",
              position = position_jitterdodge(jitter.width  = 0.15,
                                              dodge.width   = 0.5),
              size = 2, alpha = 0.7,
              show.legend = FALSE) +
  scale_fill_manual(labels = c("Conventional",
                               "Ecological"),
                    values = c("#FFA040", "#008000")) +
  scale_y_continuous(lim = c(0, 21), 
                     breaks = seq(0, 20, 5),
                     minor_breaks = seq(0, 20, by = 2) ) +
  scale_x_discrete(labels= c("105dpc", "8dpp")) +
  labs(x = "",
       y = "Calcium (serum) [mg/dl]",
       fill = "Husbandry") +
  my_theme +
  theme(legend.position = "top") +
  guides(y = guide_axis(minor.ticks = TRUE),
         fill = guide_legend(override.aes = list(linetype = 0, size  = 5)))
```

```{r, fig.height=8, fig.width=8, warning=FALSE}
plot.calc
```

# How to cite R

"All analyses were performed using R Statistical Software (version 4.4.2; R Core Team 2024)".

Reference: R Core Team (2024). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

```{r, cite-r}
citation()
version$version.string
```

```{r, cite-packages}
citation("tidyverse")
citation("kableExtra")
citation("patchwork")
citation("lmerTest")
citation("car")
citation("emmeans")
citation("performance")
citation("Hmisc")
citation("viridis")
```

# Session Info

```{r, session}
sessionInfo()
```
